Election Preferences in 2022 based on Gender and Education

### 2024-04-19 paper/paper.qmd
wrote a bunch of the paper

library(boot)
library(broom.mixed)
library(collapse)
library(dataverse)
library(gutenbergr)
library(janitor)
library(knitr)
library(marginaleffects)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)
library(arrow)
```

@@

@@

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{gender}_i + \beta_2 \times \mbox{education}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\end{align}

@@

Each game can only result in a win or a loss so there is no other outcome that needs to be considered. For the inputs, we grouped the data into several categories as it is generally accepted that 100 cs in 10 minutes is a great creep score, anything above 70 is acceptable and anything under 50 is awful. For the max level lead we stick to using values from (<-2), -1, 0, 1, (>2) as a 2 level means there is an extremely large disparity between the strength of the champions so much so that a 3rd level seems insignificant. (You can think of it like Lebron James playing against the average 10th grader vs playing against the average 7th grader.)

@@

Our results are summarized in @tbl-modelresults. This table includes values for the intercept under various conditions as well as their error value.(the value in parentheses below each intercept.) Each value of the intercept indicates the estimated log-odds of the outocme variable. What this means is that the log of the outcome is given as the intercept. Therefore by taking the exponential function we can predict the likelihood of a certain outcome given a certain variable. For instance, we can see the intercept for Post-grads were -0.94. taking the exponential function of this gives a result of 0.39. This means that a random post-grad is 0.39 times as likely to vote for trump as they are for Biden. We are not concerned with the second half of the table.

@@

first_model <-
  readRDS(file = here::here("models/first_model2022.rds"))

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

##LEAGUE OF LEGENDS 
###PLAYER ROLES

@@

##LANES

@@

##LANE MINIONS

The lane minions work in interesting ways. Firstly, experience (exp) is gained if an enemy minion dies close to you. Secondly, if you are the player that deals the killing blow (last hit) to an enemy minion, you gain a small injection of gold. At first glance, the amount seems insignificant, but it adds up over time, and usually 20 more last hits than your lane opponent grants you a gold lead that is equivalent to killing them. Usually, the first concept beginner players looking to improve are told to work on is their last hitting. Not every minion is always worth last hitting, as sometimes, in the act of trying to last hit a minion, you expose yourself to deadly attacks from the enemy. A stronger enemy can use the threat of a deadly attack to prevent their lane opponent from getting last hits. In the top lane, it is also possible to deny your lane opponent exp by positioning very aggressively; however, this usually results in you not being able to secure all the last hits for yourself. The reason this doesn’t work in mid is due to the shape of the map, in which the effective length of the lane is much shorter as towers are close to each other. Since the top lane is longer, one mistake from an unsuspecting opponent can usually lead to a kill, whereas in the mid lane, they lose health but usually not enough to kill them before they return to the safety of their tower. The bot lane, while the same length as the top lane, isn’t usually able to zone the enemy from the exp range of lane minions as there are 2 players in each bot lane, so even with a strength lead, if you are less coordinated, you can lose the fight, and gold is usually worth more to a carry than exp. However, the bot lane has other options in which they can deny the enemy exp and gold. One way of doing this would be to stack a large minion wave and crash it under the tower. (No explanation of how to do this is provided as it is unnecessary, but essentially it means you have the lane minions from 3 or 4 different waves all under the enemy tower.) Then, they try to kill the enemy bot lane that is attempting to last hit the minions under the tower. Even if the tower kills you in the process, it is a net positive as you kill them so it evens out your death, but you also deny them 3-4 waves worth of gold and exp. Usually, when pro teams do this, they get assistance from their jungler to help kill under the tower. If they play it well and in the right situation, sometimes the enemy will die without trading back a kill.

##RESOURCES

@@

##STRATEGY

@@

##Implications of Findings

@@

@@

###Weakenesses

@@

###Next Steps

@@

analysis_data <- arrow::read_parquet(here::here("data/analysis_data/ces2022clean.parquet"))

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")

>>>>

library(boot)
library(broom.mixed)
library(collapse)
library(dataverse)
library(gutenbergr)
library(janitor)
library(knitr)
library(marginaleffects)
library(modelsummary)
library(rstanarm)
library(tidybayes)
library(tidyverse)
library(arrow)
library(ggplot2)
```

@@

It is worthwhile mentioning that there are many third party data sources with the data already cleaned and presented in easily digestible formats however the accuracy of these can be called into question as many times they are incoherent statistics among them. Therefore the most reliable dataset is from the source itself.

## Variables of Interest

There are 3 variables that we are concerned with, the player's creep score(cs) at 10 minutes into the game, the player's highest level lead against his opponent, and the outcome of the game. The implications of these are discussed thoroughly in the discussion but the overarching belief is that the stronger you are, the easier it is to win the game. The higher the cs at 10 minutes is,  the stronger you are as you have more gold to buy items. The same goes for the level disparity between you and your opponent. However, being stronger does not guarentee a win. Some players may be better at converting their leads into victories and therefore we analyze one of the richest rivalries in league of legends history, T1 vs Generation Gaming.

```{r}
#| echo: false
#| eval: true
#| label: fig-t1levelunfiltered
#| fig-cap: "T1's CS at 10"
#| warning: false
t1_data <- read_parquet(here::here("data/analysis_data/t1_data.parquet"))
ggplot(t1_data, aes(x = `Max Level Lead`, fill = Outcome)) +
  geom_bar(position = "dodge") +
  labs(x = "Max Level Lead", y = "Count", title = "T1's Win Rate with Different Creep Scores") +
  scale_fill_manual(values = c("0" = "#FF6666", "1" = "#74ADD1"),
                    labels = c("Losses", "Wins")) + # Colors for Outcome 0 and 1
  theme_minimal()

@fig-t1levelunfiltered shows the wins and losses based on different level leads from all the T1 players.

```{r}
#| echo: false
#| eval: true
#| label: fig-t1level
#| fig-cap: "T1's Laner Level Lead and Wins"
#| warning: false
t1_data_filtered <- read_parquet(here::here("data/analysis_data/t1_data_filtered.parquet"))
ggplot(t1_data_filtered, aes(x = `Max Level Lead`, fill = Outcome)) +
  geom_bar(position = "dodge") +
  labs(x = "Max Level Lead", y = "Count", title = "T1's Laner's Win Rate with Different Creep Scores") +
  scale_fill_manual(values = c("0" = "#FF6666", "1" = "#74ADD1"),
                    labels = c("Losses", "Wins")) + # Colors for Outcome 0 and 1
  theme_minimal()

@fig-t1level Shows the level and win rates of the T1 players without considering their support and jungler. These are 2 of the 5 players that do not benefit from killing creeps(increasing their creep score) and their level disparity is less significant. As such, this can skew the data if we include them in our analysis. We can see this by comparing these two figures, in which the proportion of wins to losses based on different level leads changes in the way one would expect, increases for higher level leads and decreases for lower level leads. These figures serve two purposes, showing how the win rate increases with increased experience advantages(level lead) and they also show why we should exclude both support and jungler from our analysis.

## Data Summary

@fig-gengcs shows the wins and losses of all geng laners with their highest level lead while @fig-t1cs shows the same for the laners of T1. It is interesting to observe the subtle difference at the lower cs levels as these indicate differing play styles however this is not the topic of interest.

Finally, we should mention that the games of the players on each team were considered together to gain a better understanding of how the team functions rather than the individual players on the team.

```{r}
#| echo: false
#| eval: true
#| label: fig-gengcs
#| fig-cap: "Gen G's Laner's Win Rate with Different Creep Scores"
#| warning: false

geng_data_filtered <- read_parquet(here::here("data/analysis_data/geng_data_filtered.parquet"))
ggplot(geng_data_filtered, aes(x = `CS at 10`, fill = Outcome)) +
  geom_bar(position = "dodge") +
  labs(x = "CS at 10", y = "Count", title = "Gen G CS and Wins") +
  scale_fill_manual(values = c("0" = "#B19CD9", "1" = "#90EE90"),
                    labels = c("Losses", "Wins")) + # Colors for Outcome 0 and 1
  scale_x_discrete(labels = c("1" = "<50", "2" = "50-70", "3" = "71-100", "4" = ">100"))+
  theme_minimal()

```{r}
#| echo: false
#| eval: true
#| label: fig-t1cs
#| fig-cap: "T1's Laner's Win Rate with Different Creep Scores"
#| warning: false

t1_data_filtered <- read_parquet(here::here("data/analysis_data/t1_data_filtered.parquet"))
ggplot(t1_data_filtered, aes(x = `CS at 10`, fill = Outcome)) +
  geom_bar(position = "dodge") +
  labs(x = "CS at 10", y = "Count", title = "T1 CS and Wins") +
  scale_fill_manual(values = c("0" = "#B19CD9", "1" = "#90EE90"),
                    labels = c("Losses", "Wins")) + # Colors for Outcome 0 and 1
  scale_x_discrete(labels = c("1" = "<50", "2" = "50-70", "3" = "71-100", "4" = ">100"))+
  theme_minimal()

```
\newpage

@@

\begin{align} 
y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{CS}_i + \beta_2 \times \mbox{Level}_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
\end{align}

@@

Each game can only result in a win or a loss so there is no other outcome that needs to be considered. For the inputs, we grouped the data into several categories as it is generally accepted that 100 cs in 10 minutes is a great creep score, anything above 70 is acceptable and anything under 50 is awful. For the max level lead we have not obtained any value greater than 5 as this is a massive level lead. The minimum for this is 0 as both players always start at level 1.

Another advantage of using a logistic regression model is that it does not require the independence of trials. This is great as burnout is a very real thing when playing league of legends. Players frequently go on both win streaks and loss streaks which would indicate that your recent games can impact your current one. This could easily make some of these trials dependent on some of each other but this won't affect the usability of the logistic regression model. We use the default priors in this model as default priors aren't very informative and therefore let the data about the games drive the model without much interference.

@@

Our results are summarized in @tbl-modelresultst1 and @tbl-modelresultsgeng. This table includes values for the intercept under various conditions as well as their error value.(The value in parentheses below each intercept.) Each value of the intercept indicates the estimated log-odds of the outocme variable. What this means is that the log of the outcome is given as the intercept. Therefore by taking the exponential function we can predict the likelihood of a certain outcome given a certain variable. For instance, we can see the intercept for T1's Max Level Lead of 2 being 1.37. Taking the exponential function of this gives a result of 4.0149. This means that in a random game, a T1 player is 4 times as likely to win than to lose if they have a level lead of 2.

The error value gives the standard deviation of these intercepts and we notice that this gets really high for the high extremes of level leads. This is probably because these level leads are an extremely rare occurrence and there aren't many games in which the model can use to accurately predict the outcome of a game with an extremely high level lead. In the dataset, both teams won all games with a level lead of 5 so even with a high standard deviation, the model predicts that T1 players are over a million times more likely to win a game where there are 5 levels ahead of their opponent.(A 5 level lead is massive, think Lebron James vs a 10th grader)

In comparing the two, the one that stands out the most is the max level lead of 2, in which Gen G are predicted to be 10 times as likely to win than lose whereas T1 is only 4 times as likely to win. This suggests that T1 is a lot worse at converting their leads into wins. The other intercept that stands out is the CS at 10'2 which refers to having a creep score between 50 and 70 at the 10 minute mark. The model predicts in this situation, T1 is 1.5 times as likely to win than lose whereas geng is predicted to be 4 times as likely to lose than to win. This likely indicates that T1 are better at playing chaotic games and do not need to have gold to create advantages. The second halves of the tables are not of concern.

@@

t1_model <-
  readRDS(file = here::here("models/t1_model.rds"))

@@

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresultst1
#| tbl-cap: "Logistic Regression Model Predicting Likelihood of T1 Winning based on Early Indicators"
#| warning: false

modelsummary::modelsummary(
  list(
    "T1 model" = t1_model
  ),
  statistic = "mad",
  fmt = 2
)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresultsgeng
#| tbl-cap: "Logistic Regression Model Predicting Likelihood of GenG Winning based on Early Indicators"
#| warning: false

geng_model <-
  readRDS(file = here::here("models/geng_model.rds"))
modelsummary::modelsummary(
  list(
    "Gen G model" = geng_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

## League of Legends
### Player Roles

@@

### Lanes

@@

### Lane Minions 
The lane minions work in interesting ways. Firstly, experience (exp) is gained if an enemy minion dies close to you. Secondly, if you are the player that deals the killing blow (last hit) to an enemy minion, you gain a small injection of gold. At first glance, the amount seems insignificant, but it adds up over time, and usually 20 more last hits than your lane opponent grants you a gold lead that is equivalent to killing them. Usually, the first concept beginner players looking to improve are told to work on is their last hitting. Not every minion is always worth last hitting, as sometimes, in the act of trying to last hit a minion, you expose yourself to deadly attacks from the enemy. A stronger enemy can use the threat of a deadly attack to prevent their lane opponent from getting last hits. In the top lane, it is also possible to deny your lane opponent exp by positioning very aggressively; however, this usually results in you not being able to secure all the last hits for yourself. The reason this doesn’t work in mid is due to the shape of the map, in which the effective length of the lane is much shorter as towers are close to each other. Since the top lane is longer, one mistake from an unsuspecting opponent can usually lead to a kill, whereas in the mid lane, they lose health but usually not enough to kill them before they return to the safety of their tower. The bot lane, while the same length as the top lane, isn’t usually able to zone the enemy from the exp range of lane minions as there are 2 players in each bot lane, so even with a strength lead, if you are less coordinated, you can lose the fight, and gold is usually worth more to a carry than exp. However, the bot lane has other options in which they can deny the enemy exp and gold. One way of doing this would be to stack a large minion wave and crash it under the tower. (No explanation of how to do this is provided as it is unnecessary, but essentially it means you have the lane minions from 3 or 4 different waves all under the enemy tower.) Then, they try to kill the enemy bot lane that is attempting to last hit the minions under the tower. Even if the tower kills you in the process, it is a net positive as you kill them so it evens out your death, but you also deny them 3-4 waves worth of gold and exp. Usually, when pro teams do this, they get assistance from their jungler to help kill under the tower. If they play it well and in the right situation, sometimes the enemy will die without trading back a kill.

### Resources and How to Obtain Them

@@

### Strategy

@@

## Implications of Findings

@@

The model predicts Gen G to be way more likely to win a game with a 2 level lead against a lane opponent compared to T1. This is arguably the most significant type of level lead as a 1 level lead can come due to random fluctations. For instance, if both players need to kill 7 minions to level up once, if you kill the 7th minion 1 second before your opponent, you technically had a 1 level lead. A 2 level lead is really significant as it means you denied your opponent massive amounts of exp and are much stronger than them. While it is true that a 3 level lead is even more powerful, it is so much more uncommon, which can be observed in @fig-t1levelunfiltered. This argument extends further to 4 level and 5 level. This disparity means that T1 is probably less clinical in the way in which they finish out games. They are unable to convert a significant level lead into a win at the same rate as Gen G can.

The flipside of this is that the model predicts T1 to be much more likely to win a game in which they have a creep score of 50-70 at 10 minutes. This can be explained by various factors. Firstly, this level of cs is considered low for good players, which means these games probably had lots of early fighting in which you are required to outplay your opponent as you have no gold or exp advantage yet. Maybe they are more skilled and therefore better at these early fights or maybe they know the ins and outs of the champions and therefore they know which champions can win early fights against other champions. This can also be explained by Gen G having a player called "Chovy" who is known for his insanely high cs levels. If a game occurs where he is not farming at these levels, it may be unfamiliar territory and therefore difficult to play. While it is inconclusive as to whether these are due to stylistic differences or sheer skill, it is clear that the players of two of the world's best teams don't receive the same benefit from the same advantage.

@@

### Weakenesses

@@

### Next Steps

@@

analysis_data <- arrow::read_parquet(here::here("data/analysis_data/t1_data_filtered_max_level.parquet"))

@@

pp_check(t1_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(t1_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

In @fig-ppcheckandposteriorvspriorgeng-1 we implement a posterior predictive check. This shows the outcome variable with simulated variable from the posterior distribution.

In @fig-ppcheckandposteriorvspriorgeng-2 we compare the posterior with the prior. This shows how much the estimates change once the data is taken into account.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvspriorgeng
#| layout-ncol: 2
#| fig-cap: "Examining how the Gen G model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

genganalysis_data <- arrow::read_parquet(here::here("data/analysis_data/geng_data_filtered_max_level.parquet"))

genganalysis_data <- 
  genganalysis_data |> 
  slice_sample(n = 1000)

pp_check(geng_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(t1_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
## Diagnostics

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for the T1 model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(t1_model, "trace")

plot(t1_model, "rhat")

@fig-stanareyouokaygeng-1 is a trace plot. It shows lines that appear to bounce around and are horizontal. This suggests nothing wrong with the model we used.

@fig-stanareyouokaygeng-2 is a Rhat plot. It shows every value close to 1 and nothing greater than 1.1 This suggests the model we used is fine.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokaygeng
#| fig-cap: "Checking the convergence of the MCMC algorithm for the GenG model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(geng_model, "trace")

plot(geng_model, "rhat")
